---
title: "Coursera_Reproducible_Research_Assignment1"
author: "Nitish"
date: "December 27, 2016"
output: 
  html_document: 
    keep_md: yes
---
Introduction :- This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

Various questions asked in the assignemnet have been answered through the help of embedded 'r' code in this R Markdown which can be processed by knitr into HTML file.


Code for reading in the dataset and/or processing the data
```{r}
Read_ActivityData <- read.csv('activity.csv')
```
Calculate the total number of steps taken per day
```{r}
TotalStepsPerDay <- tapply(Read_ActivityData$steps, Read_ActivityData$date, sum, na.rm = TRUE)
TotalStepsPerDay
```
Histogram of the total number of steps taken each day 
```{r}
library(ggplot2)
#binwidth set to 200
qplot(TotalStepsPerDay, xlab='Total Steps Per Day', ylab = 'Frequency', binwidth = 200)
```

Mean and median number of steps taken each day
```{r}
Mean_TotalStepsPerDay <- mean(TotalStepsPerDay, na.rm = TRUE)
Median_TotalStepsPerDay <- median(TotalStepsPerDay, na.rm = TRUE)
Mean_TotalStepsPerDay
Median_TotalStepsPerDay

#To be reported :-
#Mean of Total Steps Per Day is 9354.23
#Median of Total Steps Per Day is 10395
```

Time series plot of the average number of steps taken
```{r}
#Part1
library(dplyr)

interval <- Read_ActivityData %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))

ggplot(interval, aes(x=interval, y=steps)) + geom_line(color = "red") + xlab("5-minute interval") + ylab("Average Number of steps taken")
```
 
The 5-minute interval that, on average, contains the maximum number of steps 
```{r}
interval[which.max(interval$steps),]
```
 

Report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(Read_ActivityData$steps))
```
Code to describe and show a strategy for imputing missing data and creating new dataset where missing values has been imputed
```{r}
Data1 <- Read_ActivityData
Temp1 <- is.na(Data1$steps)
Median_interval <- tapply(Data1$steps, Data1$interval, median, na.rm=TRUE, simplify=TRUE)
Data1$steps[Temp1] <- Median_interval[as.character(Data1$interval[Temp1])]
```

Checking for absence of missing values
```{r}
sum(is.na(Data1$steps))
```

Histogram of the total number of steps taken each day after missing values are imputed

```{r}
New_TotalStepsPerDay <- tapply(Data1$steps, Data1$date, sum)

library(ggplot2)
#binwidth set to 200
qplot(New_TotalStepsPerDay, xlab='Total Steps Per Day After Imputing Missing Values', ylab = 'Frequency', binwidth = 200)
```

Calculate and report the mean and median total number of steps taken per day
```{r}
Mean_New_TotalStepsPerDay <- mean(New_TotalStepsPerDay, na.rm = TRUE)
Median_New_TotalStepsPerDay <- median(New_TotalStepsPerDay, na.rm = TRUE)
Mean_New_TotalStepsPerDay
Median_New_TotalStepsPerDay
```

Impact of imputing missing values with median is Mean changes but Median remains same

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day

```{r}
library(lubridate)
Data1$date <- ymd(Data1$date)
Data1 <- mutate(Data1, Datetype = ifelse(weekdays(Data1$date) == "Saturday" | weekdays(Data1$date) == "Sunday", "weekends", "weekdays"))
Data1$Datetype <- as.factor(Data1$Datetype)
head(Data1)
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
Interval1 <- Data1 %>%
  group_by(interval, Datetype) %>%
  summarise(Imputed_steps = mean(steps))


g <- ggplot(Interval1, aes(x=interval, y=Imputed_steps, color = Datetype)) + 
geom_line() + facet_wrap(~Datetype, ncol = 1, nrow=2)
print(g)

```







